---
title: "Textbook Tutorial Ch 6"
author: Granger Moch and Sam Eckhardt
output: 
  html_document:
    toc: true
    number_sections: true 
    theme: flatly 
    highlight: zenburn
---

```{r}
library(car)
library(RcmdrMisc)
library(BCA)
data(CCS, package="BCA")
load("CCSLogistic.RData")

```
#Lift Chart for Estimation Sample 
```{r}
lift.chart(c("LinearCCS", "LogCCS", "MixedCCS", "MixedCCS2"), CCS[CCS$Sample=="Estimation",], 
  "Yes", .01, "cumulative", "Estimation")
```

The LogCCS model and the MixedCCS2 models perform similarly over the entire range of the sample.

#Lift Chart Using Validation Sample
```{r}
lift.chart(c("LinearCCS", "LogCCS", "MixedCCS", "MixedCCS2"), CCS[CCS$Sample=="Validation",], 
  "Yes", .01, "cumulative", "Validation")
```

The performance of the LogCCS and MixedCCS2 models is nearly identical. However, the MixedCCS2 model is preferred to the LogCCS model because it has fewer estimated coefficients. (parsimony)

#Incremental Response Rate Chart Using Validation Sample
```{r}
lift.chart(c("MixedCCS2"), CCS[CCS$Sample=="Validation",], "Yes", .01, "incremental", 
  "Validation")
```

We can see that targeting the most likely donors first will give good incremental response
rates.

