--- 
title: "<center>Bagging and Random Forests Tutorial</center>"
author: "<center>Granger Moch, Sam Eckhardt</center>" 
date: "<center>March 31, 2015</center> <hr>"
output:  
  html_document:
    toc: true
    number_sections: false 
    theme: flatly 
    highlight: zenburn
---    
<hr>

## Recap from last week

### Logistic Regression Model

```{r message=FALSE}
library(BCA)
data(CCS)
CCS$Sample <- create.samples(CCS, est = 0.50, val = 0.50, rand.seed = 1)
CCSwithsplitvar=CCS

logreg = glm(MonthGive ~ AdultAge + Age20t29 + Age20t39 + Age60pls + Age70pls + Age80pls + AveDonAmt +AveIncEA + DonPerYear + DwelValEA + EngPrmLang + FinUnivP + hh1mem + hh1t2mem + LastDonAmt + NewDonor + Region + SomeUnivP + YearsGive, family="binomial", data=CCSwithsplitvar[which(CCSwithsplitvar$Sample=="Estimation"),])
```
### Initial Tree Model

```{r message=FALSE}
library(rpart)
library(rpart.plot)
tree1CCS = rpart(MonthGive ~ AdultAge + Age20t29 + Age20t39 + Age60pls + Age70pls + Age80pls + AveDonAmt +AveIncEA + DonPerYear + DwelValEA + EngPrmLang + FinUnivP + hh1mem + hh1t2mem + LastDonAmt + NewDonor + Region + SomeUnivP + YearsGive, data=CCSwithsplitvar, cp=0.01, subset=Sample == "Estimation")

```
 

## Initial Tree

```{r}
rpart.plot(tree1CCS, type = 0, extra = 2, uniform = TRUE, fallen.leaves = FALSE)
```
 
## Remove Average donation and Last Donation Amount
```{r}
noAorLDA <- rpart(MonthGive ~ AdultAge + Age20t29 + Age20t39 + Age60pls + Age70pls + Age80pls + AveIncEA + DonPerYear + DwelValEA + EngPrmLang + FinUnivP + hh1mem + hh1t2mem + NewDonor + Region + SomeUnivP + YearsGive, data=CCSwithsplitvar,cp=0.001, subset=Sample == "Estimation")
rpart.plot(noAorLDA, type = 0, extra = 2, uniform = TRUE, fallen.leaves = FALSE)
```
 
## Identifying lowest cross validation error

```{r}
plotcp(noAorLDA)

```
 
##  Pruned tree with Cp at the lowest cross validation error
```{r}
cplowesterror=noAorLDA$cptable[which.min(noAorLDA$cptable[,"xerror"]),"CP"]
treeLowesterrornoAorLDA=prune(noAorLDA, cp=cplowesterror)
rpart.plot(treeLowesterrornoAorLDA, type = 0, extra = 2, uniform = TRUE, fallen.leaves = FALSE)
```
 
## Gains Chart: Estimation 

```{r echo=F}
par(mfrow=c(1,2))
lift.chart(c("logreg", "tree1CCS","treeLowesterrornoAorLDA"), CCSwithsplitvar[CCSwithsplitvar$Sample=="Estimation",], "Yes", .01, "cumulative", "Estimation")

lift.chart(c("logreg", "tree1CCS","treeLowesterrornoAorLDA"), CCSwithsplitvar[CCSwithsplitvar$Sample=="Validation",], "Yes", .01, "cumulative", "Validation")
```
 
## Bagging of CCS data
```{r message=FALSE}
library(randomForest)
set.seed(1)
baggingCCS = randomForest(MonthGive ~ AdultAge + Age20t29 + Age20t39 + Age60pls + Age70pls + Age80pls + AveDonAmt +AveIncEA + DonPerYear + DwelValEA + EngPrmLang + FinUnivP + hh1mem + hh1t2mem + LastDonAmt + NewDonor + Region + SomeUnivP + YearsGive, data=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],mtry=19,importance=TRUE)
baggingCCS
```
 
## Importance of predictor variables
```{r}
importance(baggingCCS)
```
 
## Importance of predictor variables
```{r}
varImpPlot(baggingCCS)
```
 
## Random Forest
```{r}
rfCCS=randomForest(MonthGive ~ AdultAge + Age20t29 + Age20t39 + Age60pls + Age70pls + Age80pls + AveDonAmt +AveIncEA + DonPerYear + DwelValEA + EngPrmLang + FinUnivP + hh1mem + hh1t2mem + LastDonAmt + NewDonor + Region + SomeUnivP + YearsGive, data=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],importance=TRUE) 

rfCCS
```
 
## Sample with rfCCS - Estimation data

```{r}
predrf = predict(rfCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="class")
predrfprob = predict(rfCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="prob")
head(predrf,2)
head(predrfprob,2)
```
 
## Example Confusion matrix implementation for rfCCS - Estimation Data
```{r message=FALSE}
actualvals=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",]$MonthGive
library(caret)
confmatrf=confusionMatrix(predrf,actualvals,positive="Yes")
confmatrf$table
```

 
## Example Confusion matrix statistics

```{r}
confmatrf$byClass
confmatrf$overall
```
 
## Gains Chart - Estimation data

* ROCR helps visualize binary classifier performance using three commands for each classifier (or four, if you counted the command to predict using model)
  * prediction, performance, and plot

```{r}
actualestim=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",]$MonthGive
library(ROCR)
# for logistic regression
estimpredglm= predict(logreg,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="response")
estimrocrpredglm=prediction(estimpredglm,actualestim)
estimrocrperfglm=performance(estimrocrpredglm,"tpr","rpp")
# for tree1CCS
estimpredtree1ccs = predict(tree1CCS, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",])
estimrocrpredtree1ccs=prediction(estimpredtree1ccs[,2],actualestim)
estimrocrperftree1ccs=performance(estimrocrpredtree1ccs,"tpr","rpp")
# for noAorLDA
estimprednoAorLDA = predict(noAorLDA, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",])
estimrocrprednoAorLDA=prediction(estimprednoAorLDA[,2],actualestim)
estimrocrperfnoAorLDA=performance(estimrocrprednoAorLDA,"tpr","rpp")
```
 
## Gains Chart - Estimation Data
```{r}
# for bagging
estimpredbaggingCCS = predict(baggingCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="prob")
estimrocrpredbaggingCCS=prediction(estimpredbaggingCCS[,2],actualestim)
estimrocrperfbaggingCCS=performance(estimrocrpredbaggingCCS,"tpr","rpp")
# for random forest
estimpredrfCCS = predict(rfCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="prob")
estimrocrpredrfCCS=prediction(estimpredrfCCS[,2],actualestim)
estimrocrperfrfCCS=performance(estimrocrpredrfCCS,"tpr","rpp")
```
 
## Gains Chart - Estimation Data
```{r echo=FALSE}
plot(estimrocrperfglm,col="cyan1",, main="Gains Chart: Estimation")
plot(estimrocrperftree1ccs,col="gold4", add=TRUE)
plot(estimrocrperfnoAorLDA,col="blue",add=TRUE)
plot(estimrocrperfbaggingCCS,col="purple",add=TRUE)
plot(estimrocrperfrfCCS,col="red",add=TRUE)
legend("bottomright",  c("LogReg","Tree1CCS","NoAorLDA","Bagging","Random Forest"), lty=c(1,1),col=c("cyan1","gold4","blue","purple","red"), horiz=FALSE)
abline(a=0,b=1,lwd=2,lty=2,col="gray") 
```
 
## Gains Chart - Validation Data
```{r echo=FALSE}
actualvalid=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",]$MonthGive
# for logistic regression
validpredglm= predict(logreg,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",],type="response")
validrocrpredglm=prediction(validpredglm,actualvalid)
validrocrperfglm=performance(validrocrpredglm,"tpr","rpp")

# for tree1CCS
validpredtree1ccs = predict(tree1CCS, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",])
validrocrpredtree1ccs=prediction(validpredtree1ccs[,2],actualvalid)
validrocrperftree1ccs=performance(validrocrpredtree1ccs,"tpr","rpp")

# for noAorLDA
validprednoAorLDA = predict(noAorLDA, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",])
validrocrprednoAorLDA=prediction(validprednoAorLDA[,2],actualvalid)
validrocrperfnoAorLDA=performance(validrocrprednoAorLDA,"tpr","rpp")

# for bagging
validpredbaggingCCS = predict(baggingCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",],type="prob")
validrocrpredbaggingCCS=prediction(validpredbaggingCCS[,2],actualvalid)
validrocrperfbaggingCCS=performance(validrocrpredbaggingCCS,"tpr","rpp")

# for random forest
validpredrfCCS = predict(rfCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",],type="prob")
validrocrpredrfCCS=prediction(validpredrfCCS[,2],actualvalid)
validrocrperfrfCCS=performance(validrocrpredrfCCS,"tpr","rpp")

# plotting all
plot(validrocrperfglm,col="cyan1",, main="Gains Chart: Validation")
plot(validrocrperftree1ccs,col="gold4", add=TRUE)
plot(validrocrperfnoAorLDA,col="blue",add=TRUE)
plot(validrocrperfbaggingCCS,col="purple",add=TRUE)
plot(validrocrperfrfCCS,col="red",add=TRUE)

legend("bottomright",  c("LogReg","Tree1CCS","NoAorLDA","Bagging","Random Forest"), lty=c(1,1),col=c("cyan1","gold4","blue","purple","red"), horiz=FALSE)
abline(a=0,b=1,lwd=2,lty=2,col="gray") 
```
 

## ROC Curves 

```{r echo=FALSE}
estimpredglm= predict(logreg,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="response")
estimrocrpredglm=prediction(estimpredglm,actualestim)
estimrocrperfglm=performance(estimrocrpredglm,"tpr","fpr")

# for tree1CCS
estimpredtree1ccs = predict(tree1CCS, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",])
estimrocrpredtree1ccs=prediction(estimpredtree1ccs[,2],actualestim)
estimrocrperftree1ccs=performance(estimrocrpredtree1ccs,"tpr","fpr")

# for noAorLDA
estimprednoAorLDA = predict(noAorLDA, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",])
estimrocrprednoAorLDA=prediction(estimprednoAorLDA[,2],actualestim)
estimrocrperfnoAorLDA=performance(estimrocrprednoAorLDA,"tpr","fpr")

# for bagging
estimpredbaggingCCS = predict(baggingCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="prob")
estimrocrpredbaggingCCS=prediction(estimpredbaggingCCS[,2],actualestim)
estimrocrperfbaggingCCS=performance(estimrocrpredbaggingCCS,"tpr","fpr")

# for random forest
estimpredrfCCS = predict(rfCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Estimation",],type="prob")
estimrocrpredrfCCS=prediction(estimpredrfCCS[,2],actualestim)
estimrocrperfrfCCS=performance(estimrocrpredrfCCS,"tpr","fpr")

par(mfrow=c(1,2))
# plotting all
plot(estimrocrperfglm,col="cyan1",, main="ROC Curves:Estimation")
plot(estimrocrperftree1ccs,col="gold4", add=TRUE)
plot(estimrocrperfnoAorLDA,col="blue",add=TRUE)
plot(estimrocrperfbaggingCCS,col="purple",add=TRUE)
plot(estimrocrperfrfCCS,col="red",add=TRUE)

legend("bottomright",  c("LogReg","Tree1CCS","NoAorLDA","Bagging","Random Forest"), lty=c(1,1),col=c("cyan1","gold4","blue","purple","red"), horiz=FALSE)
abline(a=0,b=1,lwd=2,lty=2,col="gray") 

## On Validation Data

actualvalid=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",]$MonthGive
# for logistic regression
validpredglm= predict(logreg,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",],type="response")
validrocrpredglm=prediction(validpredglm,actualvalid)
validrocrperfglm=performance(validrocrpredglm,"tpr","fpr")

# for tree1CCS
validpredtree1ccs = predict(tree1CCS, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",])
validrocrpredtree1ccs=prediction(validpredtree1ccs[,2],actualvalid)
validrocrperftree1ccs=performance(validrocrpredtree1ccs,"tpr","fpr")

# for noAorLDA
validprednoAorLDA = predict(noAorLDA, type="prob", newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",])
validrocrprednoAorLDA=prediction(validprednoAorLDA[,2],actualvalid)
validrocrperfnoAorLDA=performance(validrocrprednoAorLDA,"tpr","fpr")

# for bagging
validpredbaggingCCS = predict(baggingCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",],type="prob")
validrocrpredbaggingCCS=prediction(validpredbaggingCCS[,2],actualvalid)
validrocrperfbaggingCCS=performance(validrocrpredbaggingCCS,"tpr","fpr")

# for random forest
validpredrfCCS = predict(rfCCS,newdata=CCSwithsplitvar[CCSwithsplitvar$Sample == "Validation",],type="prob")
validrocrpredrfCCS=prediction(validpredrfCCS[,2],actualvalid)
validrocrperfrfCCS=performance(validrocrpredrfCCS,"tpr","fpr")

# plotting all
plot(validrocrperfglm,col="cyan1",, main="ROC Curves: Validation")
plot(validrocrperftree1ccs,col="gold4", add=TRUE)
plot(validrocrperfnoAorLDA,col="blue",add=TRUE)
plot(validrocrperfbaggingCCS,col="purple",add=TRUE)
plot(validrocrperfrfCCS,col="red",add=TRUE)

legend("bottomright",  c("LogReg","Tree1CCS","NoAorLDA","Bagging","Random Forest"), lty=c(1,1),col=c("cyan1","gold4","blue","purple","red"), horiz=FALSE)
abline(a=0,b=1,lwd=2,lty=2,col="gray")
```

