---
title: "Chapter 10 and 11 Tutorial: Cluster Analysis" 
author: Granger Moch and Sam Eckhardt
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
---
<hr>
#Chapter 10: Ward’s Method of Cluster Analysis and Principal Components

##Load Packages
```{r, message=FALSE, warning=FALSE}
library(Rcmdr)
library(car)
library(RcmdrMisc)
library(BCA)
library(knitr)
library(rgl)
library(BCA)
library(flexclust)

opts_chunk$set(fig.width=6, fig.height=6)
```

##Load Athletic Dataset
```{r}
data(Athletic, package="BCA")
```

##Explore Dataset Structure
```{r}
str(Athletic)
```

##Create Ward's Method Clustering Solution
```{r}
WardsAthletic <- hclust(dist(model.matrix(~-1 + Attnd+Fem+Finan+Grad+Teams+Violat+Win, 
                                          Athletic))^2 , method= "ward")
```

##Plot Dendrogram
```{r}
plot(WardsAthletic, main= "Cluster Dendrogram for Solution WardsAthletic", xlab= 
       "Observation Number in Data Set Athletic", 
     sub="Method=ward; Distance=squared-euclidian")
```

##Summarize Solution
###Cluster Size
```{r}
summary(as.factor(cutree(WardsAthletic, k = 4)))
```
<center>This shows the number of individuals assigned to each cluster.</center>

##Cluster Centroids
```{r}
by(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams + Violat + Win, Athletic), 
   as.factor(cutree(WardsAthletic, k = 4)), colMeans) # Cluster Centroids
```

<center>Most important variables as identified by each cluster:

Cluster 1: Winning <br>
Cluster 2: Graduation Rates<br>
Cluster 3: NCAA Rule Violations<br>
Cluster 4: Financial Success </center><br>


##Add Variable to Identify Clusters 
```{r}
Athletic$Wards4 <- assignCluster(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams 
                                              + Violat + Win, Athletic), Athletic, cutree(WardsAthletic, k = 4))
```
<center>This variable is created so that we can also compare cluster means graphically.</center>

###Plot of Means
```{r}
with(Athletic, plotMeans(Grad, Wards4, error.bars="conf.int", level=0.95))
```

<center>This means plot shows that cluster 2 cares most about “Grad”.</center>

##2D Bi-Plot of Clusters
```{r}
bpCent(prcomp(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams + Violat + Win, 
                           Athletic)), cutree(WardsAthletic, k = 4), data.pts = TRUE, centroids = TRUE, xlabs = 
         as.character(cutree(WardsAthletic, k = 4)))
```

##3D Bi Plot of Clusters
```{r}
bpCent3d(prcomp(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams + Violat + Win, 
                             Athletic)), cutree(WardsAthletic, k = 4), data.pts = TRUE, centroids = TRUE, xlabs = 
           as.character(cutree(WardsAthletic, k = 4)))
```
<br>
<hr>
<center>
#Chapter 11: K-Centroids Partitioning Cluster Analysis
</center><br>

##Diagnostic Box Plots of Athletic Dataset
```{r}
bootCVD(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams + Violat + Win, 
                     Athletic), k = 2:7, nboot = 100, nrep = 3, method = "kmn", col1 = 1, col2 = 2, dsname 
        = "Athletic")
```

##Assign KMeans4 Cluster to Athletic Dataset
```{r}
.cluster <- stepFlexclust(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams + 
                                         Violat + Win, Athletic), k = 4, nrep = 10, FUN = kcca, family = kccaFamily("kmeans"))
summary(.cluster)
print(.cluster@centers) # Cluster Centroids 
bpCent(prcomp(model.matrix(~-1 + Attnd + Fem + Finan + Grad + Teams + Violat + Win, 
                           Athletic)), clusters(.cluster), data.pts = TRUE, centroids = FALSE, xlabs = 
         as.character(clusters(.cluster)))
Athletic$KMeans4 <- as.factor(clusters(.cluster))
remove(.cluster)
```

##Contigency Table to Compare the 2-Cluster Solutions
```{r}
local({
  .Table <- xtabs(~KMeans4+Wards4, data=Athletic)
  cat("\nFrequency table:\n")
  print(.Table)
})
```

