---
title: <h1 style='text-align:center'>Chapter 3 Tutorial </h1>
author: <h4 style='text-align:center'>Granger Moch, Sam Eckhardt <hr>
output: 
  html_document:
    toc: true
    number_sections: true 
    theme: flatly 
    highlight: zenburn
---

<hr>
#Load Required Packages
</br>
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(Rcmdr))
suppressPackageStartupMessages(library(car))
Jack.Jill=read.csv("Jack.Jill.csv")
```
<hr>
#Data Cleaning
##Remove Households That Have No Female Head
</br>
```{r}
Jack.Jill= Jack.Jill %>%
filter(Employment!="No female head")
```
<hr>
##Set Row Names
</br>
```{r}
row.names(Jack.Jill) <- as.character(Jack.Jill$HH.ID)
Jack.Jill$HH.ID <- NULL
```
<hr>
#Creating Simple Summary Statistics Tutorial
##Summarize Variables 
</br>
```{r}
summary(Jack.Jill)
```
<hr>
##Numerical Summary 
</br>
```{r}
numSummary(Jack.Jill[,"Spending"], statistics=c("mean", "sd", "IQR", "quantiles"), 
  quantiles=c(0,.25,.5,.75,1))
```
<hr>
##Numerical Summary Summarized By Children Group
</br>
```{r}
numSummary(Jack.Jill[,"Spending"], groups=Jack.Jill$Children, statistics=c("mean", 
  "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```
<hr>
##Correlation Matrix
</br>
```{r}
cor(CCS[,c("AveDonAmt","DonPerYear","LastDonAmt","YearsGive")], use="complete")
```
<hr>
#Frequency Distributions and Histograms 
##Histogram Showing Spending On Children's Apparel
</br>
```{r}
with(Jack.Jill, Hist(Spending, scale="frequency", breaks=60, col="darkgray"))
```
<hr>
##Binning Continous Variable (Spending)
</br>
```{r}
Jack.Jill$Spend.Cat <- with(Jack.Jill, bin.var(Spending, bins=3, method='proportions',
   labels=c('Low','Medium','High')))
```
<hr>
##Frequency Distribution To Verify Equal Count Binning
</br>
```{r}
local({
  .Table <- with(Jack.Jill, table(Spend.Cat))
  cat("\ncounts:\n")
  print(.Table)
  cat("\npercentages:\n")
  print(round(100*.Table/sum(.Table), 2))
})
```
</br>

##Bar Graph of the Number of Children Present
</br>
```{r}
with(Jack.Jill, barplot(table(Children), xlab="Children", ylab="Frequency"))
```
</br>
<hr>
##Create New Factor Variable
</br>
```{r}
Jack.Jill$New.Children <- factor(Jack.Jill$Children, labels=c('1 Child',
  '2 Children','3+ Children','3+ Children'))

Jack.Jill$New.Children=factor(Jack.Jill$New.Children)
```
</br>
<hr>
##Barplot of New.Children factor
</br>
```{r}
with(Jack.Jill, barplot(table(New.Children), xlab="New.Children", ylab="Frequency"))
```
</br>
<hr>

##Contingency Table with Spend.Cat and New.Children
</br>
```{r}
local({
  .Table <- xtabs(~Spend.Cat+New.Children, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
</br>
<hr>
##Contingency Table with Spend.Cat and New.Children 
</br>
```{r}
local({
  .Table <- xtabs(~Spend.Cat+New.Children, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nColumn percentages:\n")
  print(colPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
  cat("\nExpected counts:\n")
  print(.Test$expected)
})
```
</br>
<hr>
##Contingency Table with Spend.Cat and Income
</br>
```{r}
local({
  .Table <- xtabs(~Spend.Cat+Income, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nColumn percentages:\n")
  print(colPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
  cat("\nExpected counts:\n")
  print(.Test$expected)
})
```
</br>
<hr>
## Create Properly Ordered Income Factor 
```{r}
Jack.Jill$Income <- with(Jack.Jill, factor(Income, levels=c('$0-$20k','$20k-$30k','$30k-$40k','$40k-$50k','$50k-$60k','$60k-$75k','$75k-$100k','$100k+'), ordered=TRUE))
```
</br>
<hr>
##Recreate Contingency Table Using the Reordered Income Variable
```{r}
local({
  .Table <- xtabs(~Spend.Cat+Income, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nColumn percentages:\n")
  print(colPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
  cat("\nExpected counts:\n")
  print(.Test$expected)
})
```
</br>
<hr>
##Contigency Table of Spend.Cat and Education 
```{r}
local({
  .Table <- xtabs(~Spend.Cat+Education, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nColumn percentages:\n")
  print(colPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
  cat("\nExpected counts:\n")
  print(.Test$expected)
})
```
</br>
<hr>
##Contigency Table of Spend.Cat and Age 
```{r}
local({
  .Table <- xtabs(~Spend.Cat+Age, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nColumn percentages:\n")
  print(colPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
  cat("\nExpected counts:\n")
  print(.Test$expected)
})
```
</br>
<hr>
</br>

##Contigency Table of Spend.Cat and Birth Country
```{r}
local({
  .Table <- xtabs(~Spend.Cat+BrthCntry, data=Jack.Jill)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nColumn percentages:\n")
  print(colPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
  cat("\nExpected counts:\n")
  print(.Test$expected)
})
```
