---
title: "Assignment 2- Multiple Regression" 
author: Granger Moch and Sam Eckhardt
output: 
  html_document:
    toc: true
    number_sections: true 
    theme: flatly 
    highlight: zenburn

---
```{r}
local({r <- getOption("repos"); 
       r["CRAN"] <- "http://cran.r-project.org"; options(repos=r)})

install.packages("ISLR")
library(dplyr)
library(ggplot2)
library(ISLR)
library(gridExtra)
data(Carseats)
```
#-
Summarize the variables and identify numeric and factor variables.

```{r}
summary(Carseats)
str(Carseats)
```

#- 
Plot histograms of numeric variables and bar plots for factor variables.

```{r}

p1= ggplot(Carseats, aes(x=Sales)) + geom_histogram()
p2= ggplot(Carseats, aes(x=CompPrice)) + geom_histogram()
p3= ggplot(Carseats, aes(x=Income)) + geom_histogram()
p4= ggplot(Carseats, aes(x=Advertising)) + geom_histogram()
p5= ggplot(Carseats, aes(x=Population)) + geom_histogram()
p6= ggplot(Carseats, aes(x=Price)) + geom_histogram()
p7= ggplot(Carseats, aes(x=Age)) + geom_histogram()
p8= ggplot(Carseats, aes(x=Education)) + geom_histogram()

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8)


p9= ggplot(Carseats, aes(x=Urban)) + geom_bar()
p10=ggplot(Carseats, aes(x=ShelveLoc)) + geom_bar()
p11=ggplot(Carseats, aes(x=US)) + geom_bar()

grid.arrange(p9,p10,p11, ncol=3)
```

#-
Using a combination of correlation matrix of numeric variables and scatterplot matrix or combinations of scatterplots, which numeric variables seem to have a relationship with Sales of car seats?
```{r}

Carseatsnum= Carseats %>%
  select(1:6, 8:9)

cor(Carseatsnum, use="complete.obs", method="pearson")
```
It doesn't seem as though any of the numeric variables have a strong relationship to sales of carseats.

#- 
Plot boxplots to visualize distribution of Sales across multiple levels of the three factor variables. Interpret how levels of a variable might be affecting sales.

```{r}

p12= ggplot(Carseats, aes(x=Urban, y=Sales)) + geom_boxplot()
p13= ggplot(Carseats, aes(x=US, y=Sales)) + geom_boxplot()
p14=ggplot(Carseats, aes(x=ShelveLoc, y=Sales)) + geom_boxplot()

grid.arrange(p12,p13,p14, ncol=3)
```

Good shelf location indicates higher sales, as does being based in the US.


#-
Develop a regression model to predict sales using all other variables, including factor variables. Which variables have an impact on sales? Interpret the coefficients of factor variables.

```{r}

reg= lm(Sales~CompPrice+Income+Advertising+Population+Price+ShelveLoc+Age+Education+Urban+US, data=Carseats)
summary(reg)

```

The variables that have an effect on sales are the variables which are statistically significant. This includes Competitor Price, Income, Advertising, Age, and Shelve Location. 

Interpreting the factor variables, medium shelf location means that it would increase sales by 1.95. Good shelf location increases sales by 4.85.

If the sale takes place in an urban area, sales increase by .12 compared to when a sale takes place in a non-urban area. However, this is not significant.

If the sale takes place in the United States, sales decrease by .18 compared to when a sale takes place in another country. However, this is not significant.

#-
Use the ANOVA model to identify which variables have a significant impact on sales?

```{r}
aov(reg)
summary(aov(reg))
anova(reg)
```