---
title: "Assignment 1- R Commands Exposure"
author: Sam Eckhardt and Granger Moch 
output: 
  html_document:
    toc: true
    number_sections: true 
    theme: flatly 
    highlight: zenburn
---

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(rCharts))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(Rcmdr))
```

```{r}
local({r <- getOption("repos"); 
       r["CRAN"] <- "http://cran.r-project.org"; options(repos=r)})

install.packages("nutshell")
data(births2006.smpl, package="nutshell")

??births2006.smpl
```

#-
How many variables and how many observations are there?
```{r}
summary(births2006.smpl)
numOp=count(births2006.smpl)
numOp
```

#-
Which variables are numeric and which variables are factor?

```{r} 
str(births2006.smpl)
```

#-
Please re-label different months from numbers to name of months (1: January)

```{r}
births2006.smpl$DOB_MM= factor(births2006.smpl$DOB_MM, 
labels= c('January', 'February','March','April','May', 'June', 'July', 'August', 'September', 'October', 'November', 'December'))
```

#-
Plot a bar graph of number of deliveries by month.
```{r}
deliveries_plot_month= ggplot(births2006.smpl, aes(x=DOB_MM)) + geom_bar() +xlab("Month") + ggtitle("Deliveries By Month") + ylab("Births")
deliveries_plot_month
```


#-
Please re-label different days of week (1: Sunday)
```{r}
births2006.smpl$DOB_WK= factor(births2006.smpl$DOB_WK,
levels=c('1','2','3','4','5','6','7'),
labels=c('Sunday','Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday'))
```

#-
Plot a bar graph of number of deliveries by day of week.

```{r}
deliveries_plot_day= ggplot(births2006.smpl, aes(x=DOB_WK, fill=DOB_WK, label=DOB_WK)) + geom_bar() +xlab("Day") + ggtitle("Deliveries By Day") + ylab("Births")
deliveries_plot_day
```

#-
Plot histograms of Weight gains of mother and Birth weight of baby.

```{r}
wgainHisto= ggplot(births2006.smpl, aes(x=WTGAIN)) +geom_histogram() +xlab("Weight-Gain(Pounds)") +ggtitle("Mothers' Weight Gain")
wgainHisto

birthw= ggplot(births2006.smpl,aes(x=DBWT)) +geom_histogram() +xlab("Birth Weight(Grams)") +ggtitle("Birth Weights")
birthw
```

#- 
Create two additional variables in which you bin Weight gains of mother and birth weight of baby into 4 different categories (for each) with approximately equal numbers in each category. Give the 4 bins understandable labels. Plot bar graphs for both these new variables.

```{r}
births2006.smpl$DBWT.Cat <- with(births2006.smpl, bin.var(DBWT, bins=4, method='proportions',
   labels=c('Small','Medium','Large','Extra Large')))

birthwbin= ggplot(births2006.smpl,aes(x=DBWT.Cat)) +geom_bar() +xlab("Birth Weight(Grams)") +ggtitle("Birth Weights")
birthwbin

births2006.smpl$WTGAIN.Cat <- with(births2006.smpl, bin.var(WTGAIN, bins=4, method='proportions',
   labels=c('Small Gain','Medium Gain','Large Gain','Very Large Gain')))

WTGAIN.Cat_bar= ggplot(births2006.smpl, aes(x=WTGAIN.Cat, fill=WTGAIN.Cat)) + geom_bar()
WTGAIN.Cat_bar
```

#-
Is there a relationship between Weight gains of mother (4 bins) and birth weight of baby (4 bins)?

```{r}

births2006sub= births2006.smpl %>%
filter(WTGAIN!="NA" & DBWT.Cat!="NA")


cont=table(births2006sub$WTGAIN.Cat, births2006sub$DBWT.Cat)
cont

chisq.test(cont) 
#ggplot(births2006sub, aes(x=WTGAIN, y=DBWT, color=DBWT.Cat)) +geom_point()+ geom_smooth(method=lm, color="black")

```
Judging from this p-value, there seems to be a significant relationship between these two factors.
#-
Is there a relationship between birth weight of baby (4 bins) and gender of baby?
```{r}

cont1=table(births2006sub$DBWT.Cat, births2006sub$SEX)

cont1

chisq.test(cont1)
```
Judging from this p-value, there seems to be a significant relationship between these two factors.

#-
Plot a bar graph of delivery method (DMETH_REC)
```{r}
ggplot(births2006.smpl, aes(x=DMETH_REC, fill=DMETH_REC)) + geom_bar() +scale_y_continuous(labels = comma)
```

#-
Subset the dataset by removing the “Unknown” in Delivery method

```{r}
births2006.smpl= births2006.smpl %>%
filter(DMETH_REC!="Unknown")

ggplot(births2006.smpl, aes(x=DMETH_REC, fill=DMETH_REC)) + geom_bar()+ scale_y_continuous(labels= comma)
```


a. Is there a relationship between delivery method (C-Section or Vaginal versus Day of the week)
```{r}
births2006sub= births2006.smpl %>%
filter(DMETH_REC!="Unknown")

births2006sub$DMETH_REC= factor(births2006sub$DMETH_REC)
cont2= table(births2006sub$DMETH_REC, births2006sub$DOB_WK)

cont2

chisq.test(cont2)

```

b. Plot bar graphs for number of deliveries on different days of week for both methods of deliveries.

```{r}
# subset1= births2006.smpl %>%
# select(DMETH_REC, DOB_WK) %>%
# filter(DMETH_REC!="Unknown")
# 
# subset1$DMETH_REC= factor(subset1$DMETH_REC)
# 
# birthsbardf2= table(subset1$DOB_WK,subset1$DMETH_REC)
# birthsbardf2=as.data.frame(birthsbardf2)
# 
# birthsbardf2$Var1= factor(birthsbardf2$Var1,
# levels=c('1','2','3','4','5','6','7'),
# labels=c('Sunday','Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday'))
# 
# 
# birthsbarplot1 <- nPlot(Freq ~ Var1, group = 'Var2', data = birthsbardf2, type = 'multiBarChart')
# 
# birthsbarplot1

```



c. Is there a relationship between delivery method and gender of baby? Present bar-graphs for both methods.
```{r}

cont3=table(births2006sub$DMETH_REC, births2006sub$SEX)

chisq.test(cont3)

```

d. Is there a relationship between delivery method and DPLURAL?

```{r fig.width=11, fig.height=6}

cont4= table(births2006sub$DMETH_REC, births2006sub$DPLURAL)
cont4
chisq.test(cont4)

```

