---
title: "R-Group Presentation"
author: "Granger Moch, Sam Eckhardt"
date: "April 25, 2015"
output:
  slidy_presentation:
    theme: flatly 
    highlight: pygments
    footer: "Granger Moch, Sam Eckhardt"
---


## Applied Data Mining for Business Decision Making Using R

<center>Customer and Business Analytics: Applied Data Mining for Business Decision Making Using R <br><br> by Daniel S. Putler and Robert E. Krider.</center> <br>

<b>Objectives:</b><br><br>
-Addressing business problems with advanced analytical tools <br> <br>
-Understanding of how data mining algorithms work <br> <br>
-Hands-on approach

##Predictive Modeling Techniques

```{r}
#Logistic Regression

```

##Logistic Regression
```{r}


```

## Load Required Packages and Data Set

```{r, warning=FALSE}
library(BCA)
library(MASS)
library(plyr)
library(ggplot2)
library(car)
library(RcmdrMisc)

biopsydf=biopsy
```


OVERSAMPLED?

```{r}
count(biopsydf$class=="benign")

#      x freq
#1 FALSE  241
#2  TRUE  458 
```

## Explore Data
```{r}
str(biopsydf)
```

## Set Record Names
```{r}
biopsydf$ID<- NULL
```
Since these “variables” are merely identifiers, they will
not actually be used in any analyses that we will do. We can ensure that
we will never accidentally use it as an analysis variable by setting HH.ID as
a record-name data type now. This step should ALWAYS be taken with a
new data set that has identification variables.

## Rename Variables
```{r}
biopsydf=rename(biopsydf, c("V1"="clump_thick", "V2"= "uniCellSize", "V3"="uniCellShape", "V4"="marg_ad", "V5"="epiCellSize","V6"="bareNuclei", "V7"="blandChrom", "V8"="normNucleoli", "V9"="mitoses"))

names(biopsydf)
str(biopsydf)
```

## Explore Data
```{r}
summary(biopsydf)
```

## Recode Target Variable
```{r}
biopsydf$class<- as.numeric(ifelse(biopsydf$class=="benign", 0, 1))

str(biopsydf$class)
```


## Create Sample
```{r}
biopsydf$Sample <- create.samples(biopsydf, est = 0.50, val = 0.50, rand.seed = 1)
```

str(biopsydf)
## Bin blandChrom Variable

```{r}
biopsydf$blandChrom.cat <- with(biopsydf, bin.var(blandChrom, bins=4, method='intervals', 
  labels=NULL))

```

##Plot of Means
```{r}
#create separate data set for these plots

with(biopsydf, plotMeans(class, blandChrom.cat, error.bars="se"))
```


## Bin clump_thick Variable

```{r}
biopsydf$clump_thick.cat <- with(biopsydf, bin.var(clump_thick, bins=5, method='intervals', 
  labels=NULL))

```


##Plot of Means
```{r}

with(biopsydf, plotMeans(class, clump_thick.cat, error.bars="se"))
```

## Bin clump_thick Variable

```{r}
biopsydf$marg_ad.cat <- with(biopsydf, bin.var(marg_ad, bins=4, method='intervals', 
  labels=NULL))

```

##Plot of Means
```{r}

with(biopsydf, plotMeans(class, marg_ad.cat, error.bars="se"))
```

## Correlation Matrix
```{r}
cor(biopsydf[,c("uniCellSize","marg_ad","uniCellShape","epiCellSize", "bareNuclei", "blandChrom", "normNucleoli", "mitoses", "clump_thick")], use="complete")
```
Uniform cell size and uniform cell shape appear to be correlated.

## Create logistic model
```{r}

biopsyGLM= glm(class~ clump_thick + uniCellSize + uniCellShape + marg_ad + epiCellSize + bareNuclei + blandChrom + normNucleoli + mitoses, family=binomial(logit), data=biopsydf)

summary(biopsyGLM)

biopsyGLM1= glm(class~ clump_thick + uniCellSize + marg_ad + epiCellSize + bareNuclei + blandChrom + normNucleoli + mitoses, family=binomial(logit), data=biopsydf)

summary(biopsyGLM1)

biopsyGLM2=glm(class~ clump_thick + uniCellShape + marg_ad + bareNuclei + blandChrom + normNucleoli, family=binomial(logit), data=biopsydf)
#unicellShape is now significant 

summary(biopsyGLM2)

biopsyGLM3=glm(class~ log(clump_thick) + log(uniCellShape) + log(marg_ad) + log(bareNuclei) + log(blandChrom) + log(normNucleoli), family=binomial(logit), data=biopsydf)

summary(biopsyGLM3)
```


```{r}
lift.chart(c("biopsyGLM", "biopsyGLM1", "biopsyGLM2", "biopsyGLM3" ), biopsydf[biopsydf$Sample=="Estimation",], 1, .51, "cumulative", "Estimation")
```
